---
title: 'R Exercises: Chapter 13'
author: "Noemi Carolina Guerra Montiel"
date: "2023-12"
output: pdf_document
---

*From: Introduction to Data Science: Data analysis and prediction algorithms with R by Rafael A.Irizarry*

In the previous exercise section, we wrangled data from a PDF file containing vital statistics from Puerto Rico. We did this for the month of September. Below we include code that does it for all 12 months.

```{r}
library(tidyverse)
library(lubridate)
library(purrr)
library(pdftools)
library(dslabs)

fn <- system.file("extdata", "RD-Mortality-Report_2015-18-180531.pdf",
                  package="dslabs")
dat <- map_df(str_split(pdf_text(fn), "\n"), function(s){
  s <- str_trim(s)
  header_index <- str_which(s, "2015")[1]
  tmp <- str_split(s[header_index], "\\s+", simplify = TRUE)
  month <- tmp[1]
  header <- tmp[-1]
  tail_index  <- str_which(s, "Total")
  n <- str_count(s, "\\d+")
  out <- c(1:header_index, which(n == 1), 
           which(n >= 28), tail_index:length(s))
  s[-out] |>  str_remove_all("[^\\d\\s]") |> str_trim() |>
    str_split_fixed("\\s+", n = 6) |> as_tibble() |> 
    setNames(c("day", header)) |>
    mutate(month = month, day = as.numeric(day)) |>
    pivot_longer(-c(day, month), names_to = "year", values_to = "deaths") |>
    mutate(deaths = as.numeric(deaths))
}) |>
  mutate(month = recode(month, 
                        "JAN" = 1, "FEB" = 2, "MAR" = 3, 
                        "APR" = 4, "MAY" = 5, "JUN" = 6, 
                        "JUL" = 7, "AGO" = 8, "SEP" = 9, 
                        "OCT" = 10, "NOV" = 11, "DEC" = 12)) |>
  mutate(date = make_date(year, month, day)) |>
  filter(date <= "2018-05-01")

```

1. We want to make a plot of death counts versus date. Confirm that the date variable are in fact dates and not strings.


2. Plot deaths versus date.
```{r}
measles |> ggplot(aes(x = date, y = cases)) + geom_line()
```

3. What time period is represented in these data?
- years

5. Repeat the plot but use the day of the year on the x-axis instead of date.


6. Compute the deaths per day by month.

7. Show the deaths per days for July and for September. What do you notice?

8. Compute deaths per week and make a plot.