---
title: "R Exercises: Chapter 7 "
author: "Noemi Carolina Guerra Montiel"
date: "November 2023"
output: pdf_document
---

*From: Introduction to Data Science: Data analysis and prediction algorithms with R by Rafael A.Irizarry*

1. Define variables containing the heights of males and females like this:
```{r}
library(dslabs)
male <- heights$height[heights$sex == "Male"]
female <- heights$height[heights$sex == "Female"]
```
How many measurements do we have for each?
```{r}
male_n <- length(male)
female_n <- length(female)
c(male_n, female_n)
```
2. Suppose we can’t make a plot and want to compare the distributions side by side. We can’t just list all the numbers. Instead, we will look at the percentiles. Create a five row table showing female_percentiles and male_percentiles with the 10th, 30th, 50th, 70th, & 90th percentiles for each sex. Then create a data frame with these two as columns.
```{r}
library(data.table)
categories <- c(0.1, 0.3, 0.5, 0.7, 0.9)
male_percentiles <- data.table(categories = categories, m_values = quantile(male, categories))
female_percentiles <- data.table(categories = categories, f_values = quantile(female, categories))
df <- data.frame(merge(male_percentiles, female_percentiles, by="categories"))
df
```
3. Study the following boxplots showing population sizes by country:
<center>
![boxplot_ex](imgs/boxplot-exercise-1)
</center>

Which continent has the country with the biggest population size?
**Asia**

4. What continent has the largest median population size?
**Africa**

5. What is median population size for Africa to the nearest million?
**10**

6. What proportion of countries in Europe have populations below 14 million?
a) 0.99
**b) 0.75**
c) 0.50
d) 0.25

7. If we use a log transformation, which continent shown above has the largest interquartile range?
**Asia**

8. Load the height data set and create a vector x with just the male heights:
```{r}
library(dslabs)
x <- heights$height[heights$sex == "Male"]
head(x)
```
What proportion of the data is between 69 and 72 inches (taller than 69, but shorter or equal to 72)? Hint: use a logical operator and mean.
```{r}
mean(x[x > 69 & x < 72]) 
```

