---
title: "R Exercises: Chapter 9"
author: "Noemi Carolina Guerra Montiel"
date: "2023-11"
output: pdf_document
---

*From: Introduction to Data Science: Data analysis and prediction algorithms with R by Rafael A.Irizarry*

Necessary libraries
```{r}
library(tidyverse) 
library(dslabs)
```


1. Run the following command to define the co2_wide object:
```{r}
co2_wide <- data.frame(matrix(co2, ncol = 12, byrow = TRUE)) |> 
  setNames(1:12) |>
  mutate(year = as.character(1959:1997))
co2_wide
```
Use the pivot_longer function to wrangle this into a tidy dataset. Call the column with the CO2 measurements co2 and call the month column month. Call the resulting object co2_tidy.
```{r}
co2_tidy <- co2_wide |> pivot_longer('1':'12', names_to = "month", values_to = "co2")
co2_tidy 
```

2. Plot CO2 versus month with a different curve for each year using this code:
```{r}
co2_tidy |> ggplot(aes(month, co2, color = year)) + geom_line()
```

If the expected plot is not made, it is probably because co2_tidy$month is not numeric:
```{r}
class(co2_tidy$month)
```
Rewrite your code to make sure the month column is numeric. Then make the plot.
```{r}
co2_tidy$month <- as.integer(co2_tidy$month)
co2_tidy |> ggplot(aes(month, co2, color=year)) + geom_line()
```
3. What do we learn from this plot?

a. CO2 measures increase monotonically from 1959 to 1997.
b. **CO2 measures are higher in the summer and the yearly average increased from 1959 to 1997.**
c. CO2 measures appear constant and random variability explains the differences.
d. CO2 measures do not have a seasonal trend.

4. Now load the admissions data set, which contains admission information for men and women across six majors and keep only the admitted percentage column:
```{r}
dat <- admissions |> select(-applicants)
dat
```
If we think of an observation as a major, and that each observation has two variables (men admitted percentage and women admitted percentage) then this is not tidy. Use the pivot_wider function to wrangle into tidy shape: one row for each major.
```{r}
wide_dat <- dat |> pivot_wider(names_from = gender, values_from = admitted)
wide_dat
```
5. Now we will try a more advanced wrangling challenge. We want to wrangle the admissions data so that for each major we have 4 observations: admitted_men, admitted_women, applicants_men and applicants_women. The trick we perform here is actually quite common: first use pivot_longer to generate an intermediate data frame and then pivot_wider to obtain the tidy data we want. We will go step by step in this and the next two exercises.

Use the pivot_longer function to create a tmp data.frame with a column containing the type of observation admitted or applicants. Call the new columns name and value.
```{r}
tmp <- admissions |> pivot_longer('admitted':'applicants', names_to = "name", values_to = "value")
tmp
```
6. Now you have an object tmp with columns major, gender, name and value. Note that if you combine the name and gender, we get the column names we want: admitted_men, admitted_women, applicants_men and applicants_women. Use the function unite to create a new column called column_name.
```{r}
tmp <- tmp |> unite(col = 'column_name', gender:name, sep = '_')
tmp
```


7. Now use the pivot_wider function to generate the tidy data with four variables for each major.
```{r}
tmp <- tmp |> pivot_wider(names_from = column_name, values_from = value)
tmp
```


8. Now use the pipe to write a line of code that turns admissions to the table produced in the previous exercise.
````{r}
admissions |> pivot_longer('admitted':'applicants', names_to = "name", values_to = "value") |> unite(col = 'column_name', gender:name, sep = '_') |> pivot_wider(names_from = column_name, values_from = value)
```

